%!TEX root = /Users/stevenmartell1/Documents/iSCAM-project/docs/iSCAM-guide/userGuide/usrGuide.tex
\section{Statistical functions \& probability distributions}
\begin{multicols}{2}
Many of the statistical functions commonly used in R have been written as negative log likelihoods and are in the \texttt{stats.cxx} library.  In this appendix is the documentation for the available functions in the stats.cxx library.  For the most part I have implemented the function based on the description from the R language, so it is possible to use \texttt{?function} name in R to learn more about the funciton.  Here I provide the formula, the actual code used to implement the function and a description of the variables. Note that some of the functions have been overloaded several times to deal with variables, vectors or a matrix.
\end{multicols}

\paragraph{dbeta} The beta distribution.
\[
	p(x|a,b) = - \ln(\Gamma(a+b))+(\ln(\Gamma(a))+\ln(\Gamma(b)))-(a-1)\ln(x)-(b-1)*\ln(1-x)
\]
the mean is given by $a/(a+b)$ and the variance is $\dfrac{ab}{(a+b)^2(a+b+a)}$
\begin{verbatim}
//beta distribution
dvariable dbeta(const dvariable& x, const double a, const double b)
{
	return - gammln(a+b)+(gammln(a)+gammln(b))-(a-1.)*log(x)-(b-1.)*log(1.-x);
}
\end{verbatim}

\paragraph{dgamma} The gamma distribution.
\[
 p(x|a,b) = -a \ln(b)+\ln(\Gamma(a))-(a-1)\ln(x)+bx
\]
where the mean and variance are given by $E(x) = a/b$ and $V(x) = a/b^2$.   Note that you can also compute $a$ and $b$ from $E(x)$ and $V(x)$ using the following relationships:
\[
 a=\frac{E(x)^2}{V(x)}, \quad b=\frac{E(x)}{V(x)}
\]
\noindent The following code is implemented in \texttt{stats.cxx} library:
\begin{verbatim}
//gamma
dvariable dgamma(const dvariable &x, const double a, const double b)
{
	return -a*log(b)+gammln(a)-(a-1.)*log(x)+b*x;
}
\end{verbatim}


\paragraph{dnorm} The normal distribution
\[
	p(x|\mu,\sigma) = 0.5\ln(2\pi)+\ln(\sigma)+0.5\frac{(x-\mu)^2}{\sigma^2}
\]
where the mean is $\mu$ and the variance is $\sigma^2$.
\begin{verbatim}
//normal distribution
dvariable dnorm(const dvariable& x, const double& mu, const double& std)
{
	double pi=3.141593;
	return 0.5*log(2.*pi)+log(std)+0.5*square(x-mu)/(std*std);
}
\end{verbatim}

\paragraph{dlnorm} The log normal distribution
\[
	p(x|\mu,\sigma) = 0.5\ln(2\pi)+\ln(\sigma)+\ln(x)+0.5\frac{(\ln(x)-\mu)^2}{\sigma^2}
\]
where the log mean is $\mu$ and the log variance is $\sigma^2$.
\begin{verbatim}
//log normal distribution
dvariable dlnorm(const dvariable& x, const double& mu, const double& std)
{
	double pi=3.141593;
	return 0.5*log(2.*pi)+log(std)+log(x)+square(log(x)-mu)/(2.*std*std);
}
\end{verbatim}


\section{Scripts}
\subsection{Bash script for creating the iscamArchive.zip \& .tar.gz files}
The following is a bash script that is used for archiving the source code files for \iscam\ and the scripts for the R code that is hosted on the Goole site.

%\begin{verbatim}
\begin{lstlisting}
#
#  bash script for creating the iscam archive
#
clear
echo "Archiving iscam files"

# script for creating the Tape Archive file
tar -czvf iscamArchive.$(date +%Y.%m.%d).tgz.gz Examples baranov.cxx\
 iscam.tpl iscam.dat iscam.r\
 stats.cxx iSCAMwin.txt Riscam_1.0.tar.gz

# script for creating a windoz zip file
zip -r iscamArchive.$(date +%Y.%m.%d).zip Examples baranov.cxx\
 iscam.tpl iscam.dat iscam.r\
 stats.cxx iSCAMwin.txt Riscam_1.0.tar.gz

exit 0
\end{lstlisting}
%\end{verbatim}

\section{R-code for figures and Tables}
%\begin{multicols}{2}

%		\tiny
%	\begin{alltt}
%	  \input{../iscam.R}\label{HakeDataFile}
%	\end{alltt}
%	\normalsize
%\lstset{language=R}

Here is the R-script that is used for reading in report files from ADMB.  Currently this is also compiled as an R-package.
\begin{scriptsize}
\lstinputlisting[language=R]{../../../src/R/lib/read.admb.R}
\end{scriptsize}
%\end{multicols}
